---
output:
  word_document: default
  html_document: default
---

## Project Phase 1  
## Author: Jennifer Weeks  
## BAN 502  


Libraries  
```{r, message=FALSE}
library(tidyverse) #tidyverse set of packages and functions
library(tidymodels)
library(glmnet) #for Lasso, ridge, and elastic net models 
library(GGally) #create ggcorr and ggpairs plots
library(MASS) #access to forward and backward selection algorithms
library(lmtest) #for the dw test
library(splines) #for nonlinear fitting
library(esquisse)
library(skimr)
```

## Read-In Data  
```{r, message=FALSE}
ames = read_csv("ames_student.csv")
```
## Data Structure  
```{r, results= "hide"}
str(ames) #all variables numeric
summary(ames) #no missingness
skim(ames)
```

## Data Visual Exploration  
```{r}
#esquisser(ames)
```

Findings: The following variables seem to have a visual relationship with Above_Median:
Paved_Drive, Garage_Cars, Fireplaces, Kitchen_Qual, Bedroom_AbvGr, Gr_Liv_Area, Full_Bath, TotRms_AbvGrd  

```{r, echo=FALSE}
ames2 = ames %>% dplyr::select( "Gr_Liv_Area", "Paved_Drive", "Fireplaces", "Full_Bath", "TotRms_AbvGrd", "Garage_Cars","Bedroom_AbvGr", "Kitchen_Qual", "Above_Median")
```

## Review Data Structure, No Missingness    
```{r, results= "hide"}
str(ames2) #all variables numeric
summary(ames2) #no missingness
skim(ames2)
```
## Change Character Variables    
```{r, message=FALSE}
ames2 = ames2 %>% mutate(Above_Median = as_factor(Above_Median))%>% mutate(Kitchen_Qual = as_factor(Kitchen_Qual)) %>% mutate(Paved_Drive = as_factor(Paved_Drive))
```

##Variable Correlation    
```{r}
ggpairs(ames2)
```

## Histograms, Stacked Bar Charts & Tables   

**Fireplaces** 
```{r}
ggplot(ames2, aes(x=Fireplaces)) + geom_histogram()
```

Fireplaces range from 1-4 in total   

```{r}
ggplot(ames2,aes(x=Fireplaces, fill = Above_Median)) + geom_bar() + theme_bw()
```

```{r}
t1 = table(ames2$Above_Median, ames2$Fireplaces)
prop.table(t1, margin = 2)
```

Having at least 1 Fireplace seems to increase odds of selling Above_Median.   

**Bedrooms Above Ground**  

```{r}
ggplot(ames2, aes(x=Bedroom_AbvGr)) + geom_histogram()
```

Bedrooms above ground range from 1-6 in total  

```{r}
ggplot(ames2,aes(x=Bedroom_AbvGr, fill = Above_Median)) + geom_bar() + theme_bw()
```
```{r}
t2 = table(ames2$Above_Median, ames2$Bedroom_AbvGr)
prop.table(t2, margin = 2)
```

Mixed variation in findings with 3-5 bedrooms having the best odds of selling Above_Median.   

**Full Baths**  

```{r}
ggplot(ames2, aes(x=Full_Bath)) + geom_histogram()
```

Full Bathrooms ranges from  1- 4 in total  

```{r}
ggplot(ames2,aes(x=Full_Bath, fill = Above_Median)) + geom_bar() + theme_bw()
```


```{r}
t3 = table(ames2$Above_Median, ames2$Full_Bath)
prop.table(t3, margin = 2)
```

Having 2 or more Full Baths increases odds of selling Above_Median  

**Ground Living Area**  
```{r}
ggplot(ames2, aes(x=Gr_Liv_Area)) + geom_histogram()
```

There are a few outliers once you get over 3000 sqft for Gr_Liv_Area  

```{r}
ames2 = ames2 %>% filter(Gr_Liv_Area < 3500)
ggplot(ames2, aes(x=Gr_Liv_Area)) + geom_histogram()
```

```{r}
ggplot(ames2,aes(x=Above_Median,y=Gr_Liv_Area)) + geom_boxplot()  + theme_bw()
```

There seems to be correlation with homes having a larger Ground Living Area selling Above_Median.   

**Kitchen Quality**  
```{r}
ggplot(ames2,aes(x=Kitchen_Qual, fill = Above_Median)) + geom_bar() + theme_bw()

```
```{r}
t4 = table(ames2$Above_Median, ames2$Kitchen_Qual)
prop.table(t4, margin = 2)
```

Kitchens that are Excellent and Good have show correlation with selling Above_Median.   

**Total Rooms Above Ground**  
```{r}
ggplot(ames2, aes(x=TotRms_AbvGrd)) + geom_histogram()
```

Few outliers over 10 rooms but overall fairly normal distribution  

```{r}
ggplot(ames2,aes(x=TotRms_AbvGrd, fill = Above_Median)) + geom_bar() + theme_bw()
```

```{r}
t5 = table(ames2$Above_Median, ames2$TotRms_AbvGrd)
prop.table(t5, margin = 2)
```

With some variation it seems the more total rooms above ground the higher likelihood of selling Above_Median.  

**Garage Cars**  
```{r}
ggplot(ames2, aes(x=Garage_Cars)) + geom_histogram()
```

Range from 0 - 5   

```{r}
ggplot(ames2,aes(x=Garage_Cars, fill = Above_Median)) + geom_bar() + theme_bw()
```

```{r}
t6 = table(ames2$Above_Median, ames2$Garage_Cars)
prop.table(t6, margin = 2)
```

Garages with 2-4 car spaces seem to have strong indication of selling Above_Median.  

**Paved Drive**  
```{r}
ggplot(ames2,aes(x=Paved_Drive, fill = Above_Median)) + geom_bar() + theme_bw()
```
```{r}
t7 = table(ames2$Above_Median, ames2$Paved_Drive)
prop.table(t7, margin = 2)
```

Having a Paved Driveway is a slight indicator of selling Above_Median  
